# Retrieving Data 

## Exercises 

```{r}
load('FRED_gold.rda')
names(gold) <- "TR"
head(gold, 3)
tail(gold, 3)
```

```{r}
logret <- diff(log(gold$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(tail(ret,3),6)
```

```{r}
logret_w <- apply.weekly(logret, sum)
round(head(logret_w,3), 6)
ret_w <- exp(logret_w) - 1

logret_m <- apply.monthly(logret, sum)
round(head(logret_m,3), 6)
ret_m <- exp(logret_m) - 1

logret_q <- apply.quarterly(logret, sum)
round(head(logret_q,3), 6)
ret_q <- exp(logret_q) - 1

logret_y <- apply.yearly(logret, sum)
round(tail(logret_y,3), 6)
ret_y <- exp(logret_y) - 1
round(tail(ret_y,3), 6)
```

## Quiz 1

Retrieve the data:

```{r}
dexusuk <- getSymbols("DEXUSUK", src="FRED", auto.assign=FALSE)
dexusuk <- na.omit(dexusuk)
# dexusuk <- 1/dexusuk
dexusuk <- dexusuk["1979-12-31/2017-12-31"]
names(dexusuk) <- "TR"
```

Calculate daily log returns and discrete returns:

```{r}
logret <- diff(log(dexusuk$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(head(ret,3),6)
```

Calculate weekly, monthly, quarterly, and yearly log returns and discrete returns:

```{r}
logret_w <- apply.weekly(logret, sum)
round(head(logret_w,3), 6)
ret_w <- exp(logret_w) - 1
round(head(ret_w,3), 6)

logret_m <- apply.monthly(logret, sum)
round(head(logret_m,3), 6)
ret_m <- exp(logret_m) - 1
round(head(ret_m,3), 6)

logret_q <- apply.quarterly(logret, sum)
round(tail(logret_q,3), 6)
ret_q <- exp(logret_q) - 1

logret_y <- apply.yearly(logret, sum)
round(tail(logret_y,3), 6)
ret_y <- exp(logret_y) - 1
round(tail(ret_y,3), 6)
```

# Risk Management Under Normal Distributions

## Exercises

```{r}
round(mean(logret), 8) # mu
round(sd(logret), 8)   # sigma
```

```{r}
mu <- mean(logret)
sigma <- sd(logret)

alpha <- 0.05
var <- qnorm(alpha, mu, sigma)
round(var, 8)
```

```{r}
HFvar <- 1000 * (exp(var) - 1)
round(HFvar, 1)
```

```{r}
es <- mu - sigma * dnorm(qnorm(alpha, 0, 1), 0, 1)/alpha
round(es, 6)

HFes <- 1000 * (exp(es) - 1)
round(HFes, 1)
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

rvec <- rnorm(100000, mu, sigma)
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

rvec <- sample(as.vector(logret), 100000, replace = TRUE)
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

## Quiz 2

Retrieve the data:

```{r}
dexszus <- getSymbols("DEXSZUS", src="FRED", auto.assign=FALSE)
dexszus <- na.omit(dexszus)
dexszus <- 1/dexszus
dexszus <- dexszus["1979-12-31/2017-12-31"]
names(dexszus) <- "TR"
```

Calculate daily log returns and discrete returns:

```{r}
logret <- diff(log(dexszus$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(head(ret,3),6)
```

Estimate VaR and ES at 99% confidence level.

```{r}
mu <- mean(logret)
sigma <- sd(logret)

round(mu, 6)
round(sigma, 6)

alpha <- 0.01
var <- qnorm(alpha, mu, sigma)
round(var, 6)
```

```{r}
es <- mu - sigma * dnorm(qnorm(alpha, 0, 1), 0, 1)/alpha
round(es, 6)
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

rvec <- rnorm(100000, mu, sigma)
VaR <- quantile(rvec, alpha)
ESN <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ESN, 6)
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

rvec <- sample(as.vector(logret), 100000, replace = TRUE)
VaR <- quantile(rvec, alpha)
ESR <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ESR, 6)
```

```{r}
HFesN <- 1000 * (exp(ESN) - 1)
HFesR <- 1000 * (exp(ESR) - 1)

round(HFesN, 2)
round(HFesR, 2)
```

# Risk Management Under Non-normal Distributions

```{r}
library(moments)
rvec <- as.vector(logret)
round(skewness(rvec), 2)
```

```{r}
round(kurtosis(rvec), 2)
```

```{r}
jarque.test(rvec)
```

```{r}
library(MASS)
t.fit <- fitdistr(rvec, "t")
round(t.fit$estimate, 6)
```

```{r}
alpha <- 0.05
RNGkind(sample.kind="Rounding")
set.seed(123789)

library(metRology)
rvec <- rt.scaled(100000, mean = t.fit$estimate[1], sd = t.fit$estimate[2], df = t.fit$estimate[3])
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])
round(VaR, 6)
round(ES, 6)
```

1. Simulate from the estimated student-t distribution.

```{r}
alpha <- 0.05
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
for (i in 1:10) {
    rvec <- rvec + rt.scaled(100000, mean = t.fit$estimate[1], sd = t.fit$estimate[2], df = t.fit$estimate[3])
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

2. Simulate from the empirical distribution with i.i.d. draws.

```{r}
alpha <- 0.05
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
for (i in 1:10) {
    rvec <- rvec + sample(as.vector(logret), 100000, replace = TRUE)
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

3. Simulate from the empirical distribution with block draws.

```{r}
alpha <- 0.05
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
rdat <- as.vector(logret)
posn <- seq(from = 1, to = length(rdat) - 9, by = 1)
rpos <- sample(posn, 100000, replace = TRUE)
for (i in 1:10) {
    rvec <- rvec + rdat[rpos]
    rpos <- rpos + 1
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

## Quiz 3

Retrieve the data:

```{r}
dexusuk <- getSymbols("DEXUSUK", src="FRED", auto.assign=FALSE)
dexusuk <- na.omit(dexusuk)
# dexusuk <- 1/dexusuk
dexusuk <- dexusuk["1979-12-31/2017-12-31"]
names(dexusuk) <- "TR"
```

Calculate daily log returns and discrete returns:

```{r}
logret <- diff(log(dexusuk$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(head(ret,3),6)
```

```{r}
library(moments)
rvec <- as.vector(logret)
round(skewness(rvec), 2)
```

```{r}
round(kurtosis(rvec), 2)
```

```{r}
jarque.test(rvec)
```

```{r}
library(MASS)
t.fit <- fitdistr(rvec, "t")
round(t.fit$estimate, 6)
```

```{r}
alpha <- 0.01
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rt.scaled(100000, mean = t.fit$estimate[1], sd = t.fit$estimate[2], df = t.fit$estimate[3])
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

1. Simulate from the estimated student-t distribution.

```{r}
alpha <- 0.01
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
for (i in 1:10) {
    rvec <- rvec + rt.scaled(100000, mean = t.fit$estimate[1], sd = t.fit$estimate[2], df = t.fit$estimate[3])
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

2. Simulate from the empirical distribution with i.i.d. draws.

```{r}
alpha <- 0.01
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
for (i in 1:10) {
    rvec <- rvec + sample(as.vector(logret), 100000, replace = TRUE)
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

3. Simulate from the empirical distribution with block draws.

```{r}
alpha <- 0.01
RNGkind(sample.kind="Rounding")
set.seed(123789)
library(metRology)

rvec <- rep(0, 100000)
rdat <- as.vector(logret)
posn <- seq(from = 1, to = length(rdat) - 9, by = 1)
rpos <- sample(posn, 100000, replace = TRUE)
for (i in 1:10) {
    rvec <- rvec + rdat[rpos]
    rpos <- rpos + 1
}
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

# Risk Management under Volatility Clustering

```{r}
load('FRED_gold.rda')
names(gold) <- "TR"
head(gold, 3)
tail(gold, 3)
```

```{r}
logret <- diff(log(gold$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(tail(ret,3),6)
```

```{r}
acf(logret)
```

```{r}
acf(abs(logret))
```

```{r}
library(rugarch)
garch.N <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
                      distribution.model = "norm")
fit.garch.N <- ugarchfit(spec = garch.N, data = logret)
fit.garch.N
```

```{r}
save1 <- cbind(logret, fit.garch.N@fit$sigma, fit.garch.N@fit$z)
names(save1) <- c("logret", "s", "z")
```

```{r}
acf(save1$z)
```

```{r}
acf(abs(save1$z))
```

```{r}
garch.t <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
                      distribution.model = "std")
fit.garch.t <- ugarchfit(spec = garch.t, data = logret)
fit.garch.t
```

```{r}
save1 <- cbind(logret, fit.garch.t@fit$sigma, fit.garch.t@fit$z)
names(save1) <- c("logret", "s", "z")
parm1 <- fit.garch.t@fit$coef
```

```{r}
acf(save1$z)
```

```{r}
acf(abs(save1$z))
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

boot.garch <- ugarchboot(fit.garch.t, 
             method = "Partial",
             sampling = "raw",
             n.ahead = 1,
             n.bootpred = 100000,
             solver = "solnp")

rvec <- boot.garch@fseries
alpha <- 0.05
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

```{r}
logret2008 <- logret["1980-01-01/2008-09-15"]

fit.garch2008 <- ugarchfit(spec = garch.t, data = logret2008)

RNGkind(sample.kind="Rounding")
set.seed(123789)

boot.garch2008 <- ugarchboot(fit.garch2008, 
             method = "Partial",
             sampling = "raw",
             n.ahead = 1,
             n.bootpred = 100000,
             solver = "solnp")

rvec <- boot.garch2008@fseries
alpha <- 0.05
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

```{r}
logret1987 <- logret["1980-01-01/1987-10-19"]

fit.garch1987 <- ugarchfit(spec = garch.t, data = logret1987)

RNGkind(sample.kind="Rounding")
set.seed(123789)

boot.garch1987 <- ugarchboot(fit.garch1987, 
             method = "Partial",
             sampling = "raw",
             n.ahead = 1,
             n.bootpred = 100000,
             solver = "solnp")

rvec <- boot.garch1987@fseries
alpha <- 0.05
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```

## Quiz 4

Retrieve the data:

```{r}
dexusuk <- getSymbols("DEXUSUK", src="FRED", auto.assign=FALSE)
dexusuk <- na.omit(dexusuk)
# dexusuk <- 1/dexusuk
dexusuk <- dexusuk["1979-12-31/2017-12-31"]
names(dexusuk) <- "TR"
```

Calculate daily log returns and discrete returns:

```{r}
logret <- diff(log(dexusuk$TR))[-1]
round(head(logret,3),6)

ret <- exp(logret)-1
round(head(ret,3),6)
```

```{r}
acf(logret)
```

```{r}
acf(abs(logret))
```

```{r}
garch.t <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
                      distribution.model = "std")
fit.garch.t <- ugarchfit(spec = garch.t, data = logret)
fit.garch.t
```

```{r}
RNGkind(sample.kind="Rounding")
set.seed(123789)

boot.garch <- ugarchboot(fit.garch.t, 
             method = "Partial",
             sampling = "raw",
             n.ahead = 1,
             n.bootpred = 100000,
             solver = "solnp")

rvec <- boot.garch@fseries
alpha <- 0.05
VaR <- quantile(rvec, alpha)
ES <- mean(rvec[rvec < VaR])

round(VaR, 6)
round(ES, 6)
```